---
title: "DeepLearningCausal"
output: github_document
date: "2025-04-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the tutorial for the main functions in the `DeepLearningCausal` package. 
```{r importcran, warning = FALSE, eval=FALSE}
devtools::install_github("hknd23/DeepLearningCausal",force = TRUE)
install.packages("DeepLearningCausal")
```

```{r importgh, warning = FALSE}
library(DeepLearningCausal)
```

```{r importdep, warning = FALSE}
library(SuperLearner)
set.seed(123456)
```

## Import Datasets and Model Speficication

```{r data}
data("exp_data")
data("pop_data")

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
```
## Ensemble Meta Learners

### Ensemble S Learner
```{r enslearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

slearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               SL.learners = SLlearners,
               binary.outcome = FALSE)

print(slearner_en)

plot(slearner_en)
```

### Ensemble T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r entlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

tlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "T.Learner",
               SL.learners = SLlearners,
               binary.outcome = FALSE)
print(tlearner_en)

plot(tlearner_en)
```

### Ensemble X Learner
For the X Learner, use `meta.learner.type = "X.Learner"`: 

```{r enxlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

xlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "X.Learner",
               SL.learners = SLlearners,
               binary.outcome = FALSE)
print(xlearner_en)

plot(xlearner_en)

```

## Deep Neural Meta Learners

### Deep Neural S Learner
```{r nnslearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

slearner_nn <- metalearner_deepneural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               stepmax = 1e+9,  algorithm = "rprop+",
               hidden.layer = c(2, 2), linear.output = FALSE,
               binary.outcome = FALSE)
print(slearner_nn)

plot(slearner_nn)
```

### Deep Neural T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r nntlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

tlearner_nn <- metalearner_deepneural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "T.Learner", stepmax = 1e+9, 
               hidden.layer = c(2, 2), linear.output = FALSE,
               binary.outcome = FALSE)
print(tlearner_nn)

plot(tlearner_nn)
```

### Deep Neural X Learner
For the X Learner, use `meta.learner.type = "X.Learner"`: 

```{r nntlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

xlearner_nn <- metalearner_deepneural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "X.Learner", stepmax = 1e+9, 
               hidden.layer = c(2, 2), linear.output = FALSE,
               binary.outcome = FALSE)
print(xlearner_nn)

plot(xlearner_nn)
```

## Ensemble PATT-C

```{r pattcen, warning=FALSE}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

pattc_en <- pattc_ensemble(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            compl.SL.learners = SLlearners,
            response.SL.learners = SLlearners,
            binary.outcome = FALSE, bootstrap = FALSE)
print(pattc_en)
```

```{r pattcenv, warning=FALSE}
plot(pattc_en)
```

### Bootstrapped Ensemble PATT-C

For bootstrapped PATT-C users can specify arguments `bootstrap = TRUE` and number
of iterations with `nboot = 5000` (default is 1000). 

```{r enmeta_boot, warning = FALSE}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_en_b <- pattc_ensemble(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            compl.SL.learners = SLlearners,
            response.SL.learners = SLlearners,
            binary.outcome = FALSE, 
            bootstrap = TRUE, nboot = 5000)
print(pattc_en_b)
```

## Deep Neural PATT-C

```{r nnpatt}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_nn <- pattc_deepneural(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            compl.hidden.layer = c(2, 2),
            response.hidden.layer = c(2, 2),
            compl.stepmax = 1e+09, response.stepmax = 1e+09,
            binary.outcome = FALSE)
print(pattc_nn)
```

```{r pattcnnv, warning=FALSE}
plot(pattc_nn)
```

### Bootstrapped Deep Neural PATT-C

For bootstrapped PATT-C, use `bootstrap = TRUE` and number of iterations with `nboot = 5000`. 

```{r nnmeta_boot}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_nn_b <- pattc_deepneural(response.formula = response_formula,
              exp.data = exp_data, pop.data = pop_data,
              treat.var = "strong_leader", compl.var = "compliance",
              compl.hidden.layer = c(2, 2),
              response.hidden.layer = c(2, 2),
              compl.stepmax = 1e+09, response.stepmax = 1e+09,
              binary.outcome = FALSE, 
              bootstrap = TRUE, nboot = 5000)
print(pattc_nn_b)
```

## Subgroup HTE Plots

Subgroup analyses using CATEs from meta learners and PATT-C predictions can be performed using `hte_plot`: 

```{r htes}
hte_plot(xlearner_nn, cut_points = c(20, 1, 3, 3, 5, 2, 1, 6), boot = TRUE,
         n_boot = 1000)

hte_plot(pattc_nn, cut_points = c(20, 1, 3, 3, 5, 2, 1, 6), boot = TRUE,
         n_boot = 1000)
```

