---
title: "DeepLearningCausal"
output: github_document
date: "2024-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the tutorial for the main functions in the `DeepLearningCausal` package. 

```{r importgh, warning = FALSE}
devtools::install_github("hknd23/DeepLearningCausal",force = TRUE)
library(DeepLearningCausal)
```
```{r importcran, warning = FALSE, eval=FALSE}
install.packages("DeepLearningCausal")
```

```{r importdep, warning = FALSE}
library(SuperLearner)
set.seed(123456)
```

## Import Datasets and Model Speficication

```{r data}
data("exp_data")
data("pop_data")

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
```
## Ensemble Meta Learners

### Ensemble S Learner
```{r enslearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

slearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               SL.learners = SLlearners,
               binary.outcome = FALSE)

print(slearner_en)
```

### Ensemble T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r entlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

tlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "T.Learner",
               SL.learners = SLlearners,
               binary.outcome = FALSE)
print(slearner_en)
```

```{r visualst}
library(ggplot2)
library(dplyr)

data.frame(tlearner_en$CATEs) %>% ggplot( aes(x= tlearner_en.CATEs)) +
  geom_histogram(alpha = 0.6, position = 'identity') +
  xlab("CATEs (T Learner)")+ylab("")

data.frame(tlearner_en$CATEs) %>% ggplot( aes(x= tlearner_en.CATEs)) +
  geom_histogram(alpha = 0.6, position = 'identity')+
  xlab("CATEs (S Learner)")+ylab("")
```

## Deep Neural Meta Learners

### Deep Neural S Learner
```{r nnslearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

slearner_nn <- metalearner_deepneural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               stepmax = 1e+9,  algorithm = "rprop+",
               hidden.layer = c(2, 2), linear.output = FALSE,
               binary.outcome = FALSE)
print(slearner_nn)
```

### Deep Neural T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r nntlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

tlearner_nn <- metalearner_deepneural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "T.Learner", stepmax = 1e+9, 
               hidden.layer = c(2, 2), linear.output = FALSE,
               binary.outcome = FALSE)
print(tlearner_nn)
```


```{r visualstnn}
library(ggplot2)

data.frame(slearner_nn$CATEs) %>% ggplot( aes(x= slearner_nn.CATEs)) +
  geom_histogram(alpha = 0.6, position = 'identity') +
  xlab("CATEs (T Learner)")+ylab("")

data.frame(tlearner_nn$CATEs) %>% ggplot( aes(x= tlearner_nn.CATEs)) +
  geom_histogram(alpha = 0.6, position = 'identity')+
  xlab("CATEs (S Learner)")+ylab("")
```

## Ensemble PATT-C

```{r pattcen, warning=FALSE}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

pattc_en <- pattc_ensemble(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            SL.learners = SLlearners,
            binary.outcome = FALSE, bootstrap = FALSE)
print(pattc_en)
```

```{r pattcenv, warning=FALSE}
library(ggplot2)
patt_en_preds <-  rbind(data.frame("predictions" = pattc_en$pop_counterfactual[,1],
                                   type = "Y_hat0"),
                        data.frame("predictions" = pattc_en$pop_counterfactual[,2],
                                   type = "Y_hat1"))
patt_en_preds %>%
  ggplot( aes(x = predictions, fill = type)) +
  geom_histogram(alpha = 0.6, position = 'identity')+xlab("Predicted Outcome")+ylab("")+
  theme(legend.position = "bottom")+
  theme(legend.title=element_blank())
```

### Bootstrapped Ensemble PATT-C

For bootstrapped PATT-C users can specify arguments `bootstrap = TRUE` and number
of iterations with `nboot = 5000` (default is 1000). 

```{r enmeta_boot, warning = FALSE}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_en_b <- pattc_ensemble(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            SL.learners = SLlearners,
            binary.outcome = FALSE, 
            bootstrap = TRUE, nboot = 5000)
print(pattc_en_b)
```

## Deep Neural PATT-C

```{r nnpatt}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_nn <- pattc_deepneural(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            compl.hidden.layer = c(2, 2),
            response.hidden.layer = c(2, 2),
            compl.stepmax = 1e+09, response.stepmax = 1e+09,
            binary.outcome = FALSE)
print(pattc_nn)
```

```{r pattcnnv, warning=FALSE}
library(ggplot2)
patt_nn_preds <-  rbind(data.frame("predictions" = pattc_nn$pop_counterfactual[,1],
                                   type = "Y_hat0"),
                        data.frame("predictions" = pattc_nn$pop_counterfactual[,2],
                                   type = "Y_hat1"))

patt_nn_preds %>%
  ggplot( aes(x = predictions, fill = type)) +
  geom_histogram(alpha = 0.6, position = 'identity')+
  xlab("Predicted Outcome")+ylab("")+
  theme(legend.position = "bottom")+
  theme(legend.title=element_blank())
```

### Bootstrapped Deep Neural PATT-C

For bootstrapped PATT-C, use `bootstrap = TRUE` and number of iterations with `nboot = 5000`. 

```{r nnmeta_boot}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_nn_b <- pattc_deepneural(response.formula = response_formula,
            exp.data = exp_data, pop.data = pop_data,
            treat.var = "strong_leader", compl.var = "compliance",
            compl.hidden.layer = c(2, 2),
            response.hidden.layer = c(2, 2),
            compl.stepmax = 1e+09, response.stepmax = 1e+09,
            binary.outcome = FALSE, 
            bootstrap = TRUE, nboot = 5000)
print(pattc_nn_b)
```

