---
title: "DeepLearningCausal"
output: github_document
date: "2025-11-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the tutorial for the functions in the `DeepLearningCausal`
package using the ensemble method from SuperLearner and neural network based on neuralnet.

```{r importcran, warning = FALSE, eval=FALSE}
devtools::install_github("hknd23/DeepLearningCausal",force = TRUE)
install.packages("DeepLearningCausal")
```

```{r importgh, warning = FALSE}
library(DeepLearningCausal)
```

```{r importdep, warning = FALSE}
library(SuperLearner)
set.seed(123456)
```

## Import Datasets and Model Speficication

```{r data}
data("exp_data")
data("pop_data")

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
```
## Ensemble Meta Learners

### Ensemble S Learner
```{r enslearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

slearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               SL.learners = SLlearners)

print(slearner_en)

plot(slearner_en)
```

### Ensemble T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r entlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

tlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "T.Learner",
               SL.learners = SLlearners)
print(tlearner_en)

plot(tlearner_en)
```

### Ensemble X Learner
For the X Learner, use `meta.learner.type = "X.Learner"`: 

```{r enxlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

xlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "X.Learner",
               SL.learners = SLlearners)
print(xlearner_en)

plot(xlearner_en)

```

### Ensemble R Learner
For the R Learner, use `meta.learner.type = "R.Learner"`: 

```{r enrlearner}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.xgboost", "SL.ranger", "SL.nnet","SL.glm")
set.seed(123456)

rlearner_en <- metalearner_ensemble(cov.formula = response_formula,
               data = exp_data,
               treat.var = "strong_leader",
               meta.learner.type = "R.Learner",
               SL.learners = SLlearners)
print(rlearner_en)

plot(rlearner_en)

```

## Deep Neural Meta Learners

### Deep Neural S Learner
```{r nnslearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

slearner_nn <- metalearner_neural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "S.Learner",
               stepmax = 1e+9,  algorithm = "rprop+",
               hidden.layer = c(1, 1), linear.output = FALSE)
print(slearner_nn)

plot(slearner_nn)
```

### Deep Neural T Learner
For the T Learner, use `meta.learner.type = "T.Learner"`: 

```{r nntlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

tlearner_nn <- metalearner_neural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "T.Learner", stepmax = 1e+9, 
               hidden.layer = c(1, 1), linear.output = FALSE)
print(tlearner_nn)

plot(tlearner_nn)
```

### Deep Neural X Learner
For the X Learner, use `meta.learner.type = "X.Learner"`: 

```{r nnxlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

xlearner_nn <- metalearner_neural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "X.Learner", stepmax = 1e+9, 
               hidden.layer = c(1, 1), linear.output = FALSE)
print(xlearner_nn)

plot(xlearner_nn)
```
### Deep Neural R Learner
For the R Learner, use `meta.learner.type = "R.Learner"`: 

```{r nnrlearner}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

rlearner_nn <- metalearner_neural(cov.formula = response_formula,
               data = exp_data, treat.var = "strong_leader",
               meta.learner.type = "R.Learner", stepmax = 1e+9, 
               hidden.layer = c(1, 1), linear.output = FALSE)
print(rlearner_nn)

plot(rlearner_nn)
```

## Ensemble PATT-C

```{r pattcen, warning=FALSE}
library(DeepLearningCausal)
data("exp_data")
library(SuperLearner)

response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
SLlearners = c("SL.nnet","SL.glm")
set.seed(123456)

pattc_en <- pattc_ensemble(response.formula = response_formula,
                           exp.data = exp_data, 
                           pop.data = pop_data,
                           treat.var = "strong_leader", 
                           compl.var = "compliance", 
                           compl.SL.learners = SLlearners,
                           response.SL.learners = SLlearners,
                           response.family = binomial(),
                           binary.preds = FALSE, 
                           bootstrap = TRUE,
                           nboot = 1000)
print(pattc_en)
```

```{r pattcenv, warning=FALSE}
plot(pattc_en)
```

###  Deep Neural PATT-C

For bootstrapped PATT-C, use `bootstrap = TRUE` and number of iterations with `nboot = 5000`. 

```{r nnmeta_boot}
response_formula <- support_war ~ age + female + education + income +
                    employed + job_loss + hindu + political_ideology
set.seed(123456)

pattc_nn <- pattc_neural(response.formula = response_formula,
              exp.data = exp_data, pop.data = pop_data,
              treat.var = "strong_leader", compl.var = "compliance",
              compl.hidden.layer = c(1, 1),
              response.hidden.layer = c(1, 1),
              compl.stepmax = 1e+09, response.stepmax = 1e+09, bootstrap = TRUE,
              nboot = 1000, binary.preds = FALSE)
print(pattc_nn)
```

## Subgroup HTE Plots

Subgroup analyses using CATEs from meta learners and PATT-C predictions can be performed using `hte_plot`. `cut_points` specifies the cut-off points for subgroups. If a covariate is binary, the value for its cut off should be between 0-1.  
```{r htelabels}
cuts = c(20, .5, 3, 3, .5, 2, .5, 6)

labels <- c("Age <= 33", "Age > 33", "Primary School", "Tertiary", "Unemployed",         
            "Employed", "Male", "Female",  "Minority", "Majority Religion" ,
            "Low Income",  "Middle Income",  "Job Secure",  "Job Insecurity",
            "Centrist", "Right-wing Partisan")
```
```{r enshte}
hte_plot(slearner_en, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```
```{r enrhte}
hte_plot(rlearner_en, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```
```{r nnshte}
hte_plot(slearner_nn, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```
```{r nnrhte}
hte_plot(rlearner_nn, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```

```{r hte_patten}
hte_plot(pattc_en, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```
```{r hte_pattnn}
hte_plot(pattc_nn, cut_points = cuts, custom_labels = labels, boot = TRUE,
         n_boot = 1000)
```